### 2016-06-03 Paper/paper.tex
minor updates

\section{Introduction}
\citet{Sykulski2016} present a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time. 
Figure \ref{fig: fig1} shows a sample 250-day drifter trajectory. 
Oceanographers use these time series to increase understanding of  ocean circulation patterns.
Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper makes a notable contribution to the statistical literature by illustrating how to directly embed knowledge of a physical system into a stochastic model. 
 \begin{figure}[h!]
  \centering
    \includegraphics[width=.6\textwidth]{ReplicatedFigures/fig1.pdf}
        \caption{A 250-day sample drifter trajectory with two 50-day periods highlighted in green and blue. The time period highlighted in green is not affected by eddies while the time period highlighted in blue is.
        \vmcomment{You could crop away the white space in the figure.}}
        	\label{fig: fig1}
\end{figure}

\par
The motion of parcels of water moving over changing latitudes is known to have a rotational component. 
Because the Earth has  different diameter at different latitudes, the speed of objects moving with Earth's rotation varies across latitudes.  
This induces a phenomena known as the Coriolis Effect, whereby very fast moving objects or objects observed over long periods, such as drifters, end up moving at speeds that are different than the Earth beneath them as they change latitude. 
Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations. 
These motions have known frequencies, referred to as inertial frequencies, which take on positive values in the Southern hemisphere and negative values in the Northern hemisphere corresponding to the different directions objects rotate in the two hemispheres.
 Oceanographers  are interested in detecting deviations from these frequencies, as they are thought to indicate eddies, or persistent circular wave patterns \citep{Kunze1985}.
  However, the ocean has a constant turbulent background flow which makes identifying eddies in real data difficult \citep{ Elipot2010}.
 The majority of the previous work in oceanography has focused on detecting eddies using satellite altimetry rather than drifter data (see \citep{Isern2003} or \citep{Fu2010}, for example), so there is limited work on detecting eddies from drifters.
\citet{Shoosmith2005} and \citet{Lankhorst2006} do identify eddies by finding periods of repeated looping in the drifters' trajectories with a simple autoregressive model.
 Similarly, \citet{Boebel2003} calculate the curvature of the trajectory  and then classify eddies as periods of consistent rotations in one direction. 
 \citet{Sykulski2016} assert that they are the first to identify eddies by explicitly modeling inertial oscillations within a stochastic model.  
 \par
In achieving this result, several statistical advances are made.  
The authors introduce an additive spectral domain model that has components corresponding to both the turbulent background and the inertial oscillations. 
The Mat\'{e}rn covariance commonly used in spatial statistics \citep{Gneiting2012} is applied for the first time to model the turbulent background. 
Because of its flexibility, this covariance structure is shown to have more desirable features than previously-used stochastic models. 
Inertial oscillations are modeled with a complex-valued Ornstein-Uhlenbeck (OU) \vmadd{process} \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations. 
Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood that compensates for both bias and aliasing and in allowing for non-stationarity through time-varying parameter estimates.
\par
The remainder of this report proceeds as follows.
 After a succinct review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it. 
 Section 3 presents the paper's results including analysis of a long time series with time-varying parameters and an application from a simulated model. 
We conclude the paper with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.

@@

	\subsection{Spectral Analysis}
	\label{sec: specAnalysis}
	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ index the observed time points.  
	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as sums of appropriately-weighted complex exponentials. 
	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations,
	\begin{align}
Z_{m} = \frac{1}{2T} \int_{-T}^{T}z(t)e^{-i\omega_{m}t} dt\vmadd{,} && z(t) = \sum_{m=-\infty}^{\infty}Z_{m}e^{i \omega_{m} t},
\end{align}
where $T$ indicates the period \citep{Percival1993}. 
To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
\begin{equation}
S_{z}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
\end{equation}
is often employed. 
This quantity represents the variance (or power) associated with each frequency \citep{Percival1993}. 
Power spectral densities are useful for statistical modeling and can be related to the autocovariance sequence of a time series using  Fourier transformation.  
We define the autocovariance  for $z(t)$ as $s(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $\vmadd{,} where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
\begin{align}
\label{eq: fourierPair}
S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-i\omega t}d \tau \Longleftrightarrow s_{z}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
\end{align}

@@

\subsubsection{Inertial oscillations}
We begin by discussing the model for the inertial oscillations. 
In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations\vmadd{:}
\begin{align}
\label{eq: deterOsc}
\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu,\\ \nonumber
\frac{\partial v}{\partial t} - f_{0}u &= G - cv,
\end{align}
where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind  \citep{Pollard1970}. 
Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows\vmadd{:}
\begin{align}
\label{eq:diffEqDeriv}
\nonumber
\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}\vmadd{,} \\ \nonumber
dz &= (F - c u- f_{0}v)dt + i(G - cv + f_{0}u)dt\\ \nonumber
&= f_{0}(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\ \nonumber
&= if_{0}(u + iv)dt - c(u + iv)dt + (F + iG)dt\\ 
&= (if_{0} - c)z dt + (F + iG)dt. 
\end{align}
\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$\vmadd{,}  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \ref{eq:diffEqDeriv},
 \begin{equation}
\label{eq: ouEq}
dz(t) = (i f_{0} -c) z(t) dt + A d Q(t). 
\end{equation}  
Equation \eqref{eq: ouEq}  is known as the complex-valued \vmdel{Ornstein-Uhlenbeck (}OU\vmdel{)} stochastic differential equation. 
In general,  OU processes are considered continuous-time equivalents to the AR(1) process. Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
 The complex-valued OU process is the complex representation of a bivariate OU process.  
 In the formulation of Equation \eqref{eq: ouEq}, $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
 The final term of Equation \eqref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.  The term $Q(t), t \geq 0$ is a standard complex Wiener process,  meaning $Q(t) = Q_{1}(t) + i Q_{2}(t)$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
 We remind the reader that Wiener processes are the stochastic models originally used to represent Brownian motion. 
  Such a process $\{Q(t, \omega): t \geq 0\}$ on the measure space $(\omega, Q, P)$ must satisfy
\begin{enumerate}
\item $Q(0, \omega) = 0$ almost everywhere
\item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $\tau$ for all $t, \tau > 0$
\end{enumerate}
\citep{Hida1980}.

To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation~\eqref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated \vmdel{by} \vmadd{from} the data. 
If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.

For time series modeling, we are interested in the autocovariance of this process at different time lags. To derive the autocovariance as given in Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ with time lag $\tau >0$ is defined as 
\begin{equation}
\label{eq: expec}
s(\tau) = \mathbb{E}[z(t) z^{*}(t + \tau)] = \mathbb{E}[u(t)u(t + \tau)] + \mathbb{E}[(v(t) v(t + \tau)] + i(\mathbb{E}[u(t) v(t + \tau)] -\mathbb{E}[u(t + \tau)v( t)])
\end{equation}
 \citep{DeIaco2003}. The differential equations for components $u(t)$ and $v(t)$ can be expressed in matrix form as 
\begin{align}
\left( \begin{array}{c} du(t) \\ dv(t) \end{array} \right) &= 
\left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right) \left( \begin{array}{cc} u(t) \\ v(t) \end{array} \right) + \left( \begin{array}{c} A d Q_{1}(t) \\ Ad Q_{2}(t) \end{array} \right). 
\end{align}
For a bivariate OU process, the covariance matrix at time lag $\tau > 0$ is 
\begin{align}
\label{eq: bivOU}
\boldsymbol{C}(\tau) = \left( \begin{array}{cc} \mathbb{E}[u(t + \tau)u(t)] & \mathbb{E}[u(t + \tau)v(t)] \\ 
\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}\boldsymbol{C}(0)\vmadd{,}
\end{align}
where $\boldsymbol{C}(0)$ gives the covariance matrix at the initial time, $t$ \citep{Schach1971}.  \citet{Sykulski2016} specify that at the initial time, $z(t) \sim$ N(0, $A^{2}/2c$). We also note $u(t)$ and $v(t)$ are uncorrelated at the initial time, giving $\mathbb{E}[u(t)v(t)] = \mathbb{E}[v(t) u(t)] = 0$. Thus, Equation \ref{eq: expec} simplifies to
\begin{equation}
\label{eq: s0}
s(0) = \frac{A^{2}}{2c} =  \mathbb{E}[u(t)^{2}] + \mathbb{E}[(v(t)^{2}]. 
\end{equation}
Since $u(t)$ and $v(t)$ are assumed to be the same type of process, Equation \eqref{eq: s0} implies $\mathbb{E}[u(t)^{2}] =\mathbb{E}[(v(t)^{2}] = A^{2}/4c$, which means Equation \eqref{eq: bivOU} can be expressed as 
\begin{align}
\label{eq: covDeriv}
 \boldsymbol{C}(\tau) =  \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\} \left( \begin{array}{cc} A^{2}/4c & 0 \\ 0 & A^{2}/4c \end{array} \right)  &= 
\frac{A^{2}}{4c} \exp (-c \tau) \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega_{0} \\ \sin \tau \omega_{0} & \cos \tau \omega_{0}  \end{array} \right)\vmadd{,}
\end{align}
where the second equality can be \vmdel{seen} \vmadd{obtained} by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
Finally, using the values obtained in Equation \eqref{eq: covDeriv} in Equation \eqref{eq: expec}, we obtain the autocovariance  as defined \vmdel{in} \vmadd{by} \citet{Sykulski2016}\vmadd{:}
\begin{align}
\label{eq: ouAC} \nonumber
s^{(o)}(\tau) & = \frac{A^{2}}{4c} \exp(-c\tau) 2(\cos \tau \omega_{0} + \sin \tau \omega_{0})\vmadd{,} \\
&= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
\end{align}
 The $(o)$ superscript denotes that this equation is for the OU component of the full model. Using Equation \eqref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
\label{eq:ouPSD}
S^{(o)}(\omega) &=  \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
\end{align}

@@

\begin{figure}[h!]
  \centering
    \includegraphics[width=.9\textwidth]{ReplicatedFigures/fig3.pdf}
        \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes. Top Left: The eastward component of the velocities. Top Right: The eastward and northward components of the velocities. Bottom: Periodograms of both processes on the decibel scale. \vmcomment{Label all axes.}}
        	\label{fig: fmbMat}
\end{figure}

@@

\subsubsection{Semi-parametric fitting}
\label{sec: semi}
The authors note that the fitting of Equation \eqref{eq: fullSpec} can be adapted to reflect background knowledge about the drifters \vmcomment{I don't like the phrase "fitting of Equation (17).}. 
They refer to making these adjustments as applying a semi-parametric approach. 
In their analysis, they highlight using knowledge of the physical system and avoiding artifacts from interpolation. 
Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative correlations \vmcomment{correlations between correlations? just checking.}. 
To avoid this, the authors only fit the model using data from either the positive or negative half of the spectrum depending on the sign of the inertial oscillation, which is known based on the latitude of the drifter.  
Similarly, sampling effects due to interpolation make the accuracy of high frequencies suspect, so the authors only use frequencies up to a specified threshold in fitting the model. Figure \ref{fig: fig4} illustrates a semi-parametric model fit.

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{ReplicatedFigures/fig4.pdf}
        \caption{Periodograms of the time series sections colored blue and green in Figure \ref{fig: fig1}. The black line denotes the inertial frequency.   The dotted lines on the left figure indicate the location of the eddy peak and the expected shift in the inertial frequency. The semi-parametric model fit for the frequencies used in model fitting is highlighted green on the left figure and blue on the right figure. The red dotted line indicates the model fit extended beyond the frequencies where it was fit. }
        	\label{fig: fig4}
\end{figure}

@@

\par For all two hundred trajectories, the full six-parameter model, given in Equation \eqref{eq: fullSpec}, is fit to each trajectory. 
The model is fit using both sides of the spectrum with frequencies up to 1.5 times the inertial frequency. 
The parameters are assumed constant, since this portion of the time series \vmdel{does} is fairly short. 
In Figure \ref{fig: fig5}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
We also display the individual periodograms and model fits. 
The resulting spectra consistently have peaks at zero and at the inertial frequency, which indicates that the model captures the turbulent background and the inertial oscillations accurately. 
\citet{Sykulski2016}  note that this highlights the method's robustness, since the individual periodograms show considerable variability, but the model fits are all quite consistent.

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{ReplicatedFigures/fig5.png}
        \caption{Left: The spectral density of one particle (blue) with the model fit overlaid for the portion of the spectra used to fit it (green) and extended to all frequencies (red). Right: The ensemble mean periodogram (blue) with the ensemble fit for the frequencies used to fit it (green) and extended to the entire spectrum (red). Individual trajectories' periodograms (light grey) and model fits (dark fit) are also displayed. }
        \label{fig: fig5}
\end{figure}

@@

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{ReplicatedFigures/fig6.pdf}
        \caption{Left: The trajectory of the drifter in the south Pacific ocean. Right: The periodogram and model fit for the section of the trajectory highlighted in red. The green portion of the model fit corresponds to the frequencies used in fitting and the red dotted line is an extension of the fit to other frequencies. }
        \label{fig: fig6}
\end{figure}

@@

\vmcomment{Remove URLs from References}

>>>>

\section{Introduction}
\citet{Sykulski2016} present a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time. 
Figure \ref{fig: fig1} shows a sample 250-day drifter trajectory. 
Oceanographers use these time series to increase understanding of  ocean circulation patterns.
Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper makes a notable contribution to the statistical literature by illustrating how to directly embed knowledge of a physical system into a stochastic model. 
 \begin{figure}[h!]
  \centering
    \includegraphics[width=.55\textwidth]{ReplicatedFigures/fig1.pdf}
        \caption{A 250-day sample drifter trajectory with two 50-day periods highlighted in green and blue. The time period highlighted in green is not affected by eddies while the time period highlighted in blue is.}
        	\label{fig: fig1}
\end{figure}

\par
The motion of parcels of water moving over changing latitudes is known to have a rotational component. 
Because the Earth has  different diameter at different latitudes, the speed of objects moving with Earth's rotation varies across latitudes.  
This induces a phenomena known as the Coriolis Effect, whereby very fast moving objects or objects observed over long periods, such as drifters, end up moving at speeds that are different than the Earth beneath them as they change latitude. 
Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations. 
These motions have known frequencies, referred to as inertial frequencies, which take on positive values in the Southern hemisphere and negative values in the Northern hemisphere corresponding to the different directions objects rotate in the two hemispheres.
 Oceanographers  are interested in detecting deviations from these frequencies, as they are thought to indicate eddies, or persistent circular wave patterns \citep{Kunze1985}.
  However, the ocean has a constant turbulent background flow which makes identifying eddies in real data difficult \citep{ Elipot2010}.
 The majority of the previous work in oceanography has focused on detecting eddies using satellite altimetry rather than drifter data (see \citep{Isern2003} or \citep{Fu2010}, for example), so there is limited work on detecting eddies from drifters.
\citet{Shoosmith2005} and \citet{Lankhorst2006} do identify eddies by finding periods of repeated looping in the drifters' trajectories with a simple autoregressive model.
 Similarly, \citet{Boebel2003} calculate the curvature of the trajectory  and then classify eddies as periods of consistent rotations in one direction. 
 \citet{Sykulski2016} assert that they are the first to identify eddies by explicitly modeling inertial oscillations within a stochastic model.  
 \par
In achieving this result, several statistical advances are made.  
The authors introduce an additive spectral domain model that has components corresponding to both the turbulent background and the inertial oscillations. 
The Mat\'{e}rn covariance commonly used in spatial statistics \citep{Gneiting2012} is applied for the first time to model the turbulent background. 
Because of its flexibility, this covariance structure is shown to have more desirable features than previously-used stochastic models. 
Inertial oscillations are modeled with a complex-valued Ornstein-Uhlenbeck (OU) process \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations. 
Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood that compensates for both bias and aliasing and in allowing for non-stationarity through time-varying parameter estimates.
\par
The remainder of this report proceeds as follows.
 After a succinct review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it. 
 Section 3 presents the paper's results including analysis of a long time series with time-varying parameters and an application from a simulated model. 
We conclude the paper with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.

@@

	\subsection{Spectral Analysis}
	\label{sec: specAnalysis}
	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ index the observed time points.  
	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as sums of appropriately-weighted complex exponentials. 
	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations,
	\begin{align}
Z_{m} = \frac{1}{2T} \int_{-T}^{T}z(t)e^{-i\omega_{m}t} dt,  && z(t) = \sum_{m=-\infty}^{\infty}Z_{m}e^{i \omega_{m} t},
\end{align}
where $T$ indicates the period \citep{Percival1993}. 
To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
\begin{equation}
S_{z}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
\end{equation}
is often employed. 
This quantity represents the variance (or power) associated with each frequency \citep{Percival1993}. 
Power spectral densities are useful for statistical modeling and can be related to the autocovariance sequence of a time series using  Fourier transformation.  
We define the autocovariance  for $z(t)$ as $s(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $, where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
\begin{align}
\label{eq: fourierPair}
S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-i\omega t}d \tau \Longleftrightarrow s_{z}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
\end{align}

@@

\subsubsection{Inertial oscillations}
We begin by discussing the model for the inertial oscillations. 
In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations:
\begin{align}
\label{eq: deterOsc}
\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu,\\ \nonumber
\frac{\partial v}{\partial t} - f_{0}u &= G - cv,
\end{align}
where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind  \citep{Pollard1970}. 
Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows:
\begin{align}
\label{eq:diffEqDeriv}
\nonumber
\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt}, \\ \nonumber
dz &= (F - c u- f_{0}v)dt + i(G - cv + f_{0}u)dt\\ \nonumber
&= f_{0}(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\ \nonumber
&= if_{0}(u + iv)dt - c(u + iv)dt + (F + iG)dt\\ 
&= (if_{0} - c)z dt + (F + iG)dt. 
\end{align}
\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$,  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \ref{eq:diffEqDeriv},
 \begin{equation}
\label{eq: ouEq}
dz(t) = (i f_{0} -c) z(t) dt + A d Q(t). 
\end{equation}  
Equation \eqref{eq: ouEq}  is known as the complex-valued OU stochastic differential equation. 
In general,  OU processes are considered continuous-time equivalents to the AR(1) process. Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
 The complex-valued OU process is the complex representation of a bivariate OU process.  
 In the formulation of Equation \eqref{eq: ouEq}, $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
 The final term of Equation \eqref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.  The term $Q(t), t \geq 0$ is a standard complex Wiener process,  meaning $Q(t) = Q_{1}(t) + i Q_{2}(t)$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
 We remind the reader that Wiener processes are the stochastic models originally used to represent Brownian motion. 
  Such a process $\{Q(t, \omega): t \geq 0\}$ on the measure space $(\omega, Q, P)$ must satisfy
\begin{enumerate}
\item $Q(0, \omega) = 0$ almost everywhere
\item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $\tau$ for all $t, \tau > 0$
\end{enumerate}
\citep{Hida1980}.

To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation~\eqref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated  from the data. 
If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.

For time series modeling, we are interested in the autocovariance of this process at different time lags. To derive the autocovariance as given in Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ with time lag $\tau >0$ is defined as 
\begin{equation}
\label{eq: expec}
s(\tau) = \mathbb{E}[z(t) z^{*}(t + \tau)] = \mathbb{E}[u(t)u(t + \tau)] + \mathbb{E}[(v(t) v(t + \tau)] + i(\mathbb{E}[u(t) v(t + \tau)] -\mathbb{E}[u(t + \tau)v( t)])
\end{equation}
 \citep{DeIaco2003}. The differential equations for components $u(t)$ and $v(t)$ can be expressed in matrix form as 
\begin{align}
\left( \begin{array}{c} du(t) \\ dv(t) \end{array} \right) &= 
\left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right) \left( \begin{array}{cc} u(t) \\ v(t) \end{array} \right) + \left( \begin{array}{c} A d Q_{1}(t) \\ Ad Q_{2}(t) \end{array} \right). 
\end{align}
For a bivariate OU process, the covariance matrix at time lag $\tau > 0$ is 
\begin{align}
\label{eq: bivOU}
\boldsymbol{C}(\tau) = \left( \begin{array}{cc} \mathbb{E}[u(t + \tau)u(t)] & \mathbb{E}[u(t + \tau)v(t)] \\ 
\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}\boldsymbol{C}(0),
\end{align}
where $\boldsymbol{C}(0)$ gives the covariance matrix at the initial time, $t$ \citep{Schach1971}.  \citet{Sykulski2016} specify that at the initial time, $z(t) \sim$ N(0, $A^{2}/2c$). We also note $u(t)$ and $v(t)$ are uncorrelated at the initial time, giving $\mathbb{E}[u(t)v(t)] = \mathbb{E}[v(t) u(t)] = 0$. Thus, Equation \ref{eq: expec} simplifies to
\begin{equation}
\label{eq: s0}
s(0) = \frac{A^{2}}{2c} =  \mathbb{E}[u(t)^{2}] + \mathbb{E}[(v(t)^{2}]. 
\end{equation}
Since $u(t)$ and $v(t)$ are assumed to be the same type of process, Equation \eqref{eq: s0} implies $\mathbb{E}[u(t)^{2}] =\mathbb{E}[(v(t)^{2}] = A^{2}/4c$, which means Equation \eqref{eq: bivOU} can be expressed as 
\begin{align}
\label{eq: covDeriv}
 \boldsymbol{C}(\tau) =  \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\} \left( \begin{array}{cc} A^{2}/4c & 0 \\ 0 & A^{2}/4c \end{array} \right)  &= 
\frac{A^{2}}{4c} \exp (-c \tau) \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega_{0} \\ \sin \tau \omega_{0} & \cos \tau \omega_{0}  \end{array} \right),
\end{align}
where the second equality can be obtained by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
Finally, using the values obtained in Equation \eqref{eq: covDeriv} in Equation \eqref{eq: expec}, we obtain the autocovariance  as defined by \citet{Sykulski2016}:
\begin{align}
\label{eq: ouAC} \nonumber
s^{(o)}(\tau) & = \frac{A^{2}}{4c} \exp(-c\tau) 2(\cos \tau \omega_{0} + \sin \tau \omega_{0}), \\
&= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
\end{align}
 The $(o)$ superscript denotes that this equation is for the OU component of the full model. Using Equation \eqref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
\label{eq:ouPSD}
S^{(o)}(\omega) &=  \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
\end{align}

@@

\begin{figure}[h!]
  \centering
    \includegraphics[width=.85\textwidth]{ReplicatedFigures/fig3.pdf}
        \caption{Simulated velocities obtained using FBM (blue) and Mat\'{e}rn (red)  processes. Top Left: The eastward component of the velocities. Top Right: The eastward and northward components of the velocities. Bottom: Periodograms of both processes on the decibel scale. }
        	\label{fig: fmbMat}
\end{figure}

@@

\subsubsection{Semi-parametric fitting}
\label{sec: semi}
The authors note that how the overall model is fit can be adjusted to reflect background knowledge.
They refer to doing this as applying a semi-parametric approach.  
In their analysis, they highlight using knowledge of the physical system and avoiding artifacts from interpolation. 
Specifically,  variability can be different for the eastward versus northward velocities  which creates spurious correlations between positive and negative frequencies. 
To avoid this, the authors only fit the model using data from either the positive or negative half of the spectrum depending on the sign of the inertial oscillation, which is known based on the latitude of the drifter.  
Similarly, sampling effects due to interpolation make the accuracy of high frequencies suspect, so the authors only use frequencies up to a specified threshold in fitting the model. Figure \ref{fig: fig4} illustrates a semi-parametric model fit.

\begin{figure}[h!]
  \centering
    \includegraphics[width=.95\textwidth]{ReplicatedFigures/fig4.pdf}
        \caption{Periodograms of the time series sections colored blue and green in Figure \ref{fig: fig1}. The black line denotes the inertial frequency.   The dotted lines on the left figure indicate the location of the eddy peak and the expected shift in the inertial frequency. The semi-parametric model fit for the frequencies used in model fitting is highlighted green on the left figure and blue on the right figure. The red dotted line indicates the model fit extended beyond the frequencies where it was fit. }
        	\label{fig: fig4}
\end{figure}

@@

\par For all two hundred trajectories, the full six-parameter model, given in Equation \eqref{eq: fullSpec}, is fit to each trajectory. 
The model is fit using both sides of the spectrum with frequencies up to 1.5 times the inertial frequency. 
The parameters are assumed constant, since this portion of the time series  is fairly short. 
In Figure \ref{fig: fig5}, we display the fit of the aggregate spectral model to a single trajectory from the simulation and to an ensemble of trajectories. 
We also display the individual periodograms and model fits. 
The resulting spectra consistently have peaks at zero and at the inertial frequency, which indicates that the model captures the turbulent background and the inertial oscillations accurately. 
\citet{Sykulski2016}  note that this highlights the method's robustness, since the individual periodograms show considerable variability, but the model fits are all quite consistent.

\begin{figure}[h!]
  \centering
    \includegraphics[width=.95\textwidth]{ReplicatedFigures/fig5.png}
        \caption{Left: The spectral density of one particle (blue) with the model fit overlaid for the portion of the spectra used to fit it (green) and extended to all frequencies (red). Right: The ensemble mean periodogram (blue) with the ensemble fit for the frequencies used to fit it (green) and extended to the entire spectrum (red). Individual trajectories' periodograms (light grey) and model fits (dark fit) are also displayed. }
        \label{fig: fig5}
\end{figure}

@@

\begin{figure}[h!]
  \centering
    \includegraphics[width=.95\textwidth]{ReplicatedFigures/fig6.pdf}
        \caption{Left: The trajectory of the drifter in the south Pacific ocean. Right: The periodogram and model fit for the section of the trajectory highlighted in red. The green portion of the model fit corresponds to the frequencies used in fitting and the red dotted line is an extension of the fit to other frequencies. }
        \label{fig: fig6}
\end{figure}

@@
